Laplacian envelope

### 2020-06-20 Notes/laplace.tex
changed notation sor uniform serial models

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_n.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\hspace{0.02\linewidth}\label{fig:serial_sticky}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_shorten}
  \end{myenuma}
  \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
  All unlabelled arrowshave transition probabilities equal to 1.
  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
  At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
  At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
  Once this reaches zero, we discard the old end state and work on the new end state.} \label{fig:heuristicmodels}
\end{figure}

@@

These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
All of these calculations will be performed at $f\potdep=\frac{1}{2}$, so the denominator will play no role.
We will work in units such that $r=1$ temporarily, only restoring $r$ at the end.

@@

Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
Looking at \eqref{eq:larea}, the only thing we need is differences in $\etw_i(s)$.
In fact, given that the only nonzero elements of $\enc$ are
%
\begin{equation}\label{eq:unienc}
  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<M,
  \qquad
  \enc_{ii-1}=-\frac{1}{2} \quad\text{for}\quad i>1,
\end{equation}
%
and $\eq_i = \frac{1}{M}$, we have
%
\begin{equation}\label{eq:uniareaeta}
  \hat{A}(s) = \frac{\etw_1(s) - \etw_M(s)}{M}.
\end{equation}
%

@@

The $i$'th row of this equation reads
%
\begin{equation}\label{eq:detarow}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
\end{equation}
%
with the following exceptions
%
\begin{equation}\label{eq:detabnduni}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}+1}+\delta\eta_{\frac{M}{2}-1}}{2} - (s+1)\delta\eta_{\frac{M}{2}} &= \frac{1}{s},\\
  \frac{\delta\eta_{\frac{M}{2}+2}+\delta\eta_{\frac{M}{2}}}{2} - (s+1)\delta\eta_{\frac{M}{2}+1} &= -\frac{1}{s},\\
  \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
\end{aligned}
\end{equation}
%
These equations are invariant under $\delta\eta_{M+1-i} \to -\delta\eta_i$, so these two must be equal.
Then we have
%
\begin{equation}\label{eq:detahalf}
\begin{aligned}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
  \quad\text{for}\quad 1<i<\frac{M}{2}\\
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\eta_i = B\alpha^{i-1} + C\alpha^{1-i},
  \qquad\text{where}\qquad
  \alpha^2 - 2(s+1)\alpha + 1 = 0,
\end{equation}
%
with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s},
  \qquad
  C = -\frac{2}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{2\etw_1(s)}{M}
   = \frac{2}{M}\prn{\frac{1}{s}+B+C}
   = \frac{2(\alpha^{\frac{M}{2}}-1)^2}{Ms(\alpha^M+1)}.
\end{equation}
%

In reality, $M$ can only take even integer values.
However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
We find that
%
\begin{multline}\label{eq:unienv}
  \alpha = \exp\prn{\pm\frac{2y_*}{M_*}},
  \quad
  s = 2\sinh^2\frac{y_*}{M_*},
  \qquad
  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\log\alpha(s)}}{s},
  \\ \text{where}\quad
  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
  \qquad
\end{multline}
%
This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%

When $M \gg 2$, the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1 \pm \sqrt{2s}$ and the heuristic envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:stickyenc}
  \begin{aligned}
  \enc_{12} = -\enc_{MM-1} &= \frac{\varepsilon}{2},
  &\qquad
  \eq_1 = \eq_M &= \frac{1}{2+(M-2)\varepsilon},
  \\ &&
  \eq_i &=  \frac{\varepsilon}{2+(M-2)\varepsilon}
  \quad\text{for}\quad 1<i<M.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:stickyareaeta}
  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s))}{2+(M-2)\varepsilon}
       = \frac{2\varepsilon\, \etw_1(s)}{2+(M-2)\varepsilon}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndsticky}
\begin{aligned}
  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:stickycoeffs}
\begin{aligned}
  B &= -\frac{ 2 \alpha^{\frac{M}{2}} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
  C &= -\frac{ 2 \alpha^{\frac{M}{2}} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
\end{aligned}
\end{equation}
%
which leads to the Laplace transform
%
\begin{equation}\label{eq:stickylaplace}
  \hat{A}(s) = \frac{2\varepsilon}{[2+(M-2)\varepsilon]s} \,
      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
\end{equation}
%

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=1$:
%
\begin{equation}\label{eq:stickyvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M\alpha[(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^M+1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M-2}+1)]
     \\ &+ 2(\alpha^{\frac{M}{2}}-1)^2(\alpha^M+1) = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha}.
\end{aligned}
\end{equation}
%
The ``sticky'' models are optimal at values of $s$ smaller than this.
This is not quite the same as the lower boundary of the region \eqref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
The width of this gap is insignificant for larger $M$.

For small $s \ll M^{-2}$, it is given by
%
\begin{equation}\label{eq:stickyeps}
  \varepsilon = \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
\end{equation}
%
and the Laplace transform is
%
\begin{equation}\label{eq:stickyenv}
  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:shortenenc}
  \begin{aligned}
  \enc_{M-1M} = -\enc_{21} &= \frac{(1-\varepsilon)}{2},
  &\qquad
  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{M-2\varepsilon},
  \\ &&
  \eq_i &=  \frac{1}{M-2\varepsilon}
  \quad\text{for}\quad 1<i<M.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:shortenareaeta}
  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s)) + \varepsilon(\etw_2(s)-\etw_{M-1}(s))}{M-2\varepsilon}
       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}{M-2\varepsilon}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + (1-\varepsilon)(\alpha-1))}
         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - (1-\varepsilon)\alpha(\alpha-1))}
         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [(1-\varepsilon)(\alpha^{\frac{M}{2}}-1)^2
          + \varepsilon(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $M-2$ at $\varepsilon=1$.
This is easier to see if we define
%
\begin{equation}\label{eq:shortdefs}
\begin{aligned}
  F_M(\alpha) &= (\alpha^{\frac{M}{2}} - 1)^2, &
  \Delta F_M(\alpha) &= F_M(\alpha) - \zeta F_{M-2}(\alpha), &
  f_m(\alpha) = \frac{F_M(\alpha)}{\Delta F_M(\alpha)}, \\
  G_M(\alpha) &= \alpha^{M} + 1, &
  \Delta G_M(\alpha) &= G_M(\alpha) - \zeta G_{M-2}(\alpha), &
  g_m(\alpha) = \frac{G_M(\alpha)}{\Delta G_M(\alpha)}, \\
  \text{where}\quad \zeta &= \alpha^2 - \alpha + 1, &
  m &= \frac{M}{2}. \\
\end{aligned}
\end{equation}
%
Then the uniform model \eqref{eq:unilaplace} has \( \hat{A}(s) = \frac{F_m(\alpha)}{s m G_m(\alpha)} \).
Then \cref{eq:shortenlaplace} can be be written as
%
\begin{equation}\label{eq:shortsimple}
  \hat{A}(s) =
    \frac{(1-\varepsilon)F_m(\alpha) + \varepsilon \zeta F_{m-1}(\alpha)}
         {s (m - \varepsilon) ((1-\varepsilon)G_m(\alpha) + \varepsilon \zeta G_{m-1}(\alpha))}
    % = \frac{F_m - \varepsilon \Delta F_{m}}
    %      {s (m - \varepsilon) (G_m - \varepsilon \Delta G_{m})}
    = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
    \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
\end{equation}
%

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item\label{fig:short_opt}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_opt.pdf}}
  \item\label{fig:short_ratio}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_ratio.pdf}}
\end{myenuma}
\caption{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
(\ref{fig:short_opt}): Solid grey line is \cref{eq:unienv}, the uniform model with \(M\) treated as a continuous parameter and optimised.
Dotted grey line is the approximate expression from \cref{eq:unienvapprox}.
The other curves are from \cref{eq:shortopt}, optimised shortened models for various values of \(M\).
(\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
The dotted orange line is the extension of the \(M=4\) line with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
\label{fig:shorten}}
\end{center}
\end{figure}

>>>>

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\hspace{0.02\linewidth}\label{fig:serial_sticky}
    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_shorten}
  \end{myenuma}
  \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
  All unlabelled arrowshave transition probabilities equal to 1.
  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
  At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
  At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
  Once this reaches zero, we discard the old end state and work on the new end state.} \label{fig:heuristicmodels}
\end{figure}

@@

These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
All of these calculations will be performed at $f\potdep=\frac{1}{2}$, so the denominator will play no role.
We will work in units such that $r=1$ temporarily, only restoring $r$ at the end.
We study models with \(m\) strong and \(m\) weak states for a total of  \(M = 2 m\).

@@

Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
Looking at \eqref{eq:larea}, the only thing we need is differences in $\etw_i(s)$.
In fact, given that the only nonzero elements of $\enc$ are
%
\begin{equation}\label{eq:unienc}
  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<2m,
  \qquad
  \enc_{ii-1}=-\frac{1}{2} \quad\text{for}\quad i>1,
\end{equation}
%
and $\eq_i = \frac{1}{2m}$, we have
%
\begin{equation}\label{eq:uniareaeta}
  \hat{A}(s) = \frac{\etw_1(s) - \etw_{2m}(s)}{2m}.
\end{equation}
%

@@

The $i$'th row of this equation reads
%
\begin{equation}\label{eq:detarow}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
\end{equation}
%
with the following exceptions
%
\begin{equation}\label{eq:detabnduni}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{m+1}+\delta\eta_{m-1}}{2} - (s+1)\delta\eta_{m} &= \frac{1}{s},\\
  \frac{\delta\eta_{m+2}+\delta\eta_{m}}{2} - (s+1)\delta\eta_{m+1} &= -\frac{1}{s},\\
  \frac{\delta\eta_{2m-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_{2m} &= 0.
\end{aligned}
\end{equation}
%
These equations are invariant under $\delta\eta_{2m+1-i} \to -\delta\eta_i$, so these two must be equal.
Then we have
%
\begin{equation}\label{eq:detahalf}
\begin{aligned}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
  \quad\text{for}\quad 1<i<m, \\
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\eta_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
  \qquad\text{where}\qquad
  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}},
\end{equation}
%
with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{1}{s(\e^{-\beta} + 1) \cosh \prn{m \beta}},
  \qquad
  C = -\frac{1}{s(\e^{\beta} + 1) \cosh \prn{m \beta}}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{2\etw_1(s)}{2m}
   = \frac{1}{m}\prn{\frac{1}{s}+B+C}
   = \frac{2\sinh^2\prn{\frac{m \beta}{2}}}{m s \cosh\prn{m \beta}}
   = \frac{2 S\prn{m \beta}}{M s \prn{S\prn{m \beta} + 1}}.
\end{equation}
%

In reality, $M$ can only take even integer values.
However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
We find that
%
\begin{multline}\label{eq:unienv}
  M_* = \pm\frac{2y_*}{\beta},
  \quad
  s = 2\sinh^2\frac{y_*}{M_*},
  \qquad
  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\beta}}{s},
  \\ \text{where}\quad
  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
  \qquad
\end{multline}
%
This is only valid for
%
\begin{equation}\label{eq:univalid}
  2 \leq M_* \leq M,
  \qquad\implies\qquad
  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
\end{equation}
%

When $M \gg 2$, the validity region is approximately
%
\begin{equation}\label{eq:univalidapprox}
  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
\end{equation}
%
When $s \sim \CO(M^{-2}) \ll 1$, we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
%
\begin{equation}\label{eq:unienvapprox}
  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:stickyenc}
  \begin{aligned}
  \enc_{1,2} = -\enc_{2m,2m-1} &= \frac{1-\varepsilon}{2},
  &\quad
  \eq_1 = \eq_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
  \\ &&
  \eq_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
  \quad\text{for}\quad 1<i<2m.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:stickyareaeta}
  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s))}
                    {2(m-(m-1)\varepsilon)}
       = \frac{2(1-\varepsilon)\, \etw_1(s)}{2(m-(m-1)\varepsilon)}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndsticky}
\begin{aligned}
  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{1-\varepsilon}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:stickycoeffs}
\begin{aligned}
  B &= -\frac{\e^{\beta} - \varepsilon}
      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]}, \\
  C &= -\frac{1 - \varepsilon\e^{\beta}}
      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]},
\end{aligned}
\end{equation}
%
which leads to the Laplace transform
%
\begin{equation}\label{eq:stickylaplace}
  \hat{A}(s) = \frac{(1-\varepsilon)}{[m-(m-1)\varepsilon]s} \,
      \brk{\frac{S(m\beta) - \varepsilon S((m-1)\beta)}
      {S(m\beta) - \varepsilon S((m-1)\beta) + 1 - \varepsilon}}.
\end{equation}
%

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0$:
%
\begin{equation}\label{eq:stickyvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} &\propto
    m \prn{S(m\beta) - S((m-1)\beta)} + (m-1) S(m\beta) \prn{S(m\beta) + 1} = 0,\\
   &\text{where} \quad s =  S(\beta) = 2 \sinh^2(\beta/2).
\end{aligned}
\end{equation}
%
The ``sticky'' models are optimal at values of $s$ smaller than this.
This is not quite the same as the lower boundary of the region \eqref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
The width of this gap is insignificant for larger $M$.

For small $s \ll M^{-2}$, it is given by
%
\begin{equation}\label{eq:stickyeps}
  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
\end{equation}
%
and the Laplace transform is
%
\begin{equation}\label{eq:stickyenv}
  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
\end{equation}
%

@@

The following elements of $\enc$ and $\eq$ differ from \eqref{eq:unienc}:
%
\begin{equation}\label{eq:shortenenc}
  \begin{aligned}
  \enc_{2m-1,2m} = -\enc_{2,1} &= \frac{(1-\varepsilon)}{2},
  &\qquad
  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{2(m-\varepsilon)},
  \\ &&
  \eq_i &=  \frac{1}{2(m-\varepsilon)}
  \quad\text{for}\quad 1<i<M2m.
  \end{aligned}
\end{equation}
%
This means that the Laplace transform is given by
%
\begin{equation}\label{eq:shortenareaeta}
  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_{2m}(s)) 
                      + \varepsilon(\etw_2(s)-\etw_{2m-1}(s))}
                    {2(m-\varepsilon)}
       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}
              {2(m-\varepsilon)}.
\end{equation}
%

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\e^{(i-2)\beta} + C\e^{(2-i)\beta} \qquad
        \text{for} \quad 2 \leq i \leq m, \\
  \delta\eta_1 &= \frac{1}
         {[(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]s}, \\
  B &= - \frac{(2s+1) \e^{\beta/2} + (1-\varepsilon) \tanh(\beta/2)}
         {s [(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]}, \\
  C &= - \frac{((2s+1) \e^{-\beta/2} - (1-\varepsilon) \tanh(\beta/2)}
         {s [(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{(1-\varepsilon) S(m\beta) + \varepsilon(2s+1) S((m-1)\beta)}
          { s [m-\varepsilon]
            [(1-\varepsilon)[S(m\beta) + 1] + \varepsilon(2s+1)[S((m-1)\beta) + 1]] }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $m-1$ at $\varepsilon=1$.

To find the optimal \(\varepsilon\) at any given \(s\), it helps if we define
%
\begin{equation}\label{eq:shortdefs}
\begin{aligned}
  \Delta f_m(\beta) &= S(m\beta) - \zeta S((m-1)\beta), &
  f_m(\beta) &= \frac{S(m\beta)}{\Delta f_m(\beta)}, \\
  \Delta g_m(\beta) &= \Delta f_m(\beta) + 1 - \zeta , &
  g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
  \text{where}\quad \zeta &= 2s + 1. \\
\end{aligned}
\end{equation}
%
Then \cref{eq:shortenlaplace} can be be written as
%
\begin{equation}\label{eq:shortsimple}
  \hat{A}(s) =
    = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
    \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
\end{equation}
%

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item\label{fig:short_opt}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_opt.pdf}}
  \item\label{fig:short_ratio}\aligntop{%
    \includegraphics[width=0.45\linewidth]{shorten_ratio.pdf}}
\end{myenuma}
\caption{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
(\ref{fig:short_opt}): Solid grey line is \cref{eq:unienv}, the uniform model with \(M\) treated as a continuous parameter and optimised.
Dotted grey line is the approximate expression from \cref{eq:unienvapprox}.
The other curves are from \cref{eq:shortopt}, optimised shortened models for various values of \(M\).
(\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
The dotted blue and orange lines are the extensions of the \(M=2,4\) lines with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
Noting the scale of the y-axis, we see that \cref{eq:unienv,eq:unienvapprox} do provide good approximations, especially for larger \(M\).
\label{fig:shorten}}
\end{center}
\end{figure}

We can maximise this \wrt $\varepsilon$.
%
\begin{equation}\label{eq:shortopt}
  (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
  \qquad
  \hat{A}(s) = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
    \frac{1}{\brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
\end{equation}
%
We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    m(2s+1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
    - S(m\beta) [S(m\beta) + 1]
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    (m-1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
    - (2s+1) S((m-1)\beta) [S((m-1)\beta) + 1]
      = 0,\\
    &\text{where}\quad
    s = S(\beta) = 2 \sinh^2(\beta/2),
\end{aligned}
\end{equation}
%
with the $\varepsilon=0$ case providing the lower boundary and $\varepsilon=1$ providing the upper boundary.
For lower \(s\), \(\varepsilon = 0\) is the optimum. 
For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
This is depicted by the dotted lines in \cref{fig:shorten}\ref{fig:short_ratio}.
